---
title: "The Supermarket’s Strategic Plan for the Chip Category"
author: "Jose Alberto"
date: "03/11/2021"
output: github_document
---

```{r}
rm(list=ls(all=TRUE))
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```



Load Packages:

```{r}
library(rmarkdown)
library(htmltools)
library(tidyverse)
library(dplyr)
library(plyr)
library(stringr)
library(openxlsx)
library(lubridate)
library(ggplot2)
library(GGally)
library(fpp2)
library(stats)
library(fastcluster)
library(stringdist)
```

Load datasets 

```{r}
customers_transaction_dataset <- read.csv2("datasets/QVI_transaction_data.csv",
                                           sep =";",
                                           na.strings = "")
```


```{r}
customers_purchase_dataset <- read.csv("datasets/QVI_purchase_behaviour.csv",
                                       sep = ",",
                                       na.strings = "")
```


## Data exploration 
```{r}
customers_transaction_dataset %>% glimpse()
```

```{r}
str(customers_purchase_dataset)
```

```{r}
dim(customers_transaction_dataset)
```


```{r}
dim(customers_purchase_dataset)
```

Checking missing values
```{r}
sapply(customers_transaction_dataset, function(x) sum(is.na(x)))
```


```{r}
colSums(is.na(customers_purchase_dataset))
```


### Data cleaning is the process of preparing data for analysis 

```{r}
customers_transaction_dataset$ï..DATE <-  convertToDateTime(customers_transaction_dataset$ï..DATE, origin = "1900-01-01")
```


```{r}
customers_transaction_dataset$ï..DATE <- as.Date(customers_transaction_dataset$ï..DATE, origin = "1900-01-01")
```


```{r}
class(customers_transaction_dataset$ï..DATE)
```


```{r}
Stringsextract <- function(string){
  unlist(str_extract(string,regex("\\b[A-z ]*[[:punct:]]*[[:alpha:]]* [[:upper:]]*[[:lower:]]*[[:alpha:]]*")))
}
```



```{r}
customers_transaction_dataset["PRODUCT_NAME"] <- 
  Stringsextract(customers_transaction_dataset$PROD_NAME)
```



```{r}
unique(customers_transaction_dataset$PROD_NAME)
```


```{r}
customers_transaction_dataset$PRODUCT_NAME <-  mapvalues(customers_transaction_dataset$PRODUCT_NAME, from = c(
        "Natural Chip        Compny SeaSalt" ,  
        "CCs Nacho Cheese    " ,                
        "Smiths Crinkle Cut  Chips Chicken ",   
        "Smiths Chip Thinly  S",               
        "Kettle Tortilla ChpsHny&Jlpno Chili" , 
        "Old El Paso Salsa   Dip Tomato Mild ", 
        "Smiths Crinkle Chips Salt & Vinegar",
        "Grain Waves         Sweet Chilli ",    
       "Doritos Corn Chip Mexican Jalapeno ",  
       "Grain Waves Sour    Cream&Chives ",   
       "Kettle Sensations   Siracha Lime ",   
       "Twisties Cheese     ",                
       "WW Crinkle Cut      Chicken " ,        
       "Thins Chips Light& ",                  
       "CCs Original ",                       
       "Burger Rings ",                       
       "NCC Sour Cream & ",                    
       "Doritos Corn Chip Southern Chicken ",  
       "Cheezels Cheese Box ",                 
       "Smiths Crinkle      Original ",        
       "Infzns Crn Crnchers Tangy Gcamole ",   
       "Kettle Sea Salt     And Vinegar ",     
       "Smiths Chip Thinly  Cut Original ",    
       "Kettle Original ",                    
       "Red Rock Deli Thai  Chilli&Lime ",     
       "Pringles Sthrn FriedChicken ",         
       "Pringles Sweet&Spcy BBQ",              
       "Red Rock Deli SR    Salsa & Mzzrlla",
       "Thins Chips         Originl saltd ",   
       "Red Rock Deli Sp    Salt & Truffle",   
       "Smiths Thinly       Swt Chli",         
       "Kettle Chilli ",                       
       "Doritos Mexicana    ",
       "Smiths Crinkle Cut  French OnionDip ",
       "Natural ChipCo      Hony Soy Chckn",   
       "Dorito Corn Chp     Supreme ",         
       "Twisties Chicken",
       "Smiths Thinly Cut   Roast Chicken ",   
       "Smiths Crinkle Cut  Tomato Salsa ",
       "Kettle Mozzarella   Basil & Pesto",
       "Infuzions Thai SweetChili PotatoMix ",
       "Kettle Sensations   Camembert & Fig",
       "Smith Crinkle Cut   Mac N Cheese ",
       "Kettle Honey Soy    Chicken ",
       "Thins Chips Seasonedchicken ",
       "Smiths Crinkle Cut  Salt & Vinegar",
       "Infuzions BBQ Rib   Prawn Crackers ",
       "GrnWves Plus Btroot & Chilli",
       "Tyrrells Crisps     Lightly Salted ",
       "Kettle Sweet Chilli And Sour Cream ",
       "Doritos Salsa       Medium ",
       "Kettle ",
       "Pringles SourCream  Onion ",
       "Doritos Corn Chips  Original ",
       "Twisties Cheese     Burger ",
       "Old El Paso Salsa   Dip Chnky Tom Ht",
       "Cobs Popd Swt/Chlli ",
       "Woolworths Mild     Salsa ",
       "Natural Chip Co     Tmato Hrb&Spce ",
       "Cobs Popd Sea Salt  Chips ",
       "Smiths Crinkle Cut  Chips Chs",
       "French Fries Potato Chips ",
       "Old El Paso Salsa   Dip Tomato Med ",
       "Doritos Corn Chips  Cheese Supreme ",
       "Pringles Original   Crisps ",
       "RRD Chilli& ",
       "WW Original Corn    Chips ",
       "Thins Potato Chips  Hot & Spicy",
       "Cobs Popd Sour Crm  &Chives Chips",
       "Smiths Crnkle Chip  Orgnl Big Bag ",
       "Doritos Corn Chips  Nacho Cheese ",
       "Kettle Sensations   BBQ&Maple ",
       "WW D/Style Chip",
       "Pringles Chicken    Salt Crips ",
       "WW Original Stacked Chips ",
       "Smiths Chip Thinly  CutSalt",
       "Cheezels Cheese ",
       "Tostitos Lightly    Salted ",
       "Thins Chips Salt & ",
       "Smiths Crinkle Cut  Chips Barbecue ",
       "Cheetos Puffs ",
       "RRD Sweet Chilli & ",
       "WW Crinkle Cut      Original ",
       "Tostitos Splash Of  Lime ",
       "Woolworths Medium   Salsa ",
       "Kettle Tortilla ChpsBtroot&Ricotta ",
       "CCs Tasty Cheese    ",
       "Woolworths Cheese   Rings ",
       "Tostitos Smoked     Chipotle ",
       "Pringles Barbeque   ",
       "WW Supreme Cheese   Corn Chips ",
       "Pringles Mystery    Flavour ",
       "Tyrrells Crisps     Ched & Chives",
       "Snbts Whlgrn Crisps Cheddr&Mstrd ",
       "Cheetos Chs & Bacon",
       "Pringles Slt Vingar ",
       "Infuzions SourCream&Herbs Veg",
       "Kettle Tortilla ChpsFeta&Garlic ",
       "Infuzions Mango     Chutny Papadums ",
       "RRD Steak & ",
       "RRD Honey Soy       Chicken ",
       "Sunbites Whlegrn    Crisps Frch/Onin ",
       "RRD Salt & Vinegar",
       "Doritos Cheese      Supreme ",
       "Smiths Crinkle Cut  Snag&Sauce ",
       "WW Sour Cream &OnionStacked Chips",
       "RRD Lime & Pepper",
       "Natural ChipCo Sea  Salt & Vinegr",
       "Red Rock Deli Chikn&Garlic Aioli",
       "RRD SR Slow Rst     Pork Belly ",
       "RRD Pc Sea Salt     ",
       "Smith Crinkle Cut   Bolognese ",
       "Doritos Salsa Mild  "), 
          to = c("Natural Chip SeaSalt",
                 "CCs Nacho Cheese",
                 "Smiths Crinckle Chicken Chips",
                 "Smiths Thinly SourCream Onion",
                 "Kettle Tortilla Chips Chilli Honey Jalapeno",
                 "Old El Paso Tomato Salsa Dip Mild",
                 "Smiths Crinkle Chips Salt Vinegar",
                 "GrainWaves Sweet Chilli",
                 "Doritos Corn Chip Mexican Jalapeno",
                 "GrainWaves SourCream Chives",
                 "Kettle Sensations Sriracha Lime",
                 "Twisties Cheese",
                 "WW Crinkle Cut Chicken ",
                 "Thins Chips Light Tangy",
                 "CCs Original",
                 "Burger Rings",
                 "NCC SourCream Garden Chives",
                 "Doritos Corn Chip Southern Chicken",
                 "Cheezels Cheese Box",
                 "Smiths Crinkle Original",
                 "Infuzions Corn Crunchers Tangy Guacamole",
                 "Kettle SeaSalt Vinegar",
                 "Smiths Thinly Cut Original Chips",
                 "Kettle Original",
                 "RRD Thai Chilli Lime",
                 "Pringles Southern Fried Chicken",
                 "Pringles Sweet Spicy BBQ",
                 "RRD SR Salsa Mozzarella",
                 "Thins Chips Original Salted",
                 "RRD Sp Salt Truffle",
                 "SMITHS THINLY SWEET CHILLI SOURCREAM",
                 "Kettle Chilli",
                 "Doritos Mexicana",
                 "Smiths Crinkle French Onion Dip",
                 "Natural Chip Honey Soy Chicken",
                 "Doritos Corn Chips Cheese Supreme",
                 "Twisties Chicken",
                 "Smiths Thinly Roast Chicken",
                 "Smiths Crinkle Tomato Salsa",
                 "Kettle Mozzarella Basil Pesto",
                 "infuzions Potato Mix Thai Sweet Chilli",
                 "Kettle Sensations Camembert Fig",
                 "Smith Crinkle Mac N Cheese",
                 "Kettle Honey Soy Chicken ",
                 "Thins Chips Seasoned Chicken",
                 "Smiths Crinkle Salt Vinegar",
                 "InfuZions Prawn Crackers BBQ Rib",
                 "Grainwaves Beetroot Chilli Jam",
                 "Tyrrells Crisps Lightly Salted",
                 "Kettle Sweet Chilli SourCream",
                 "Doritos Medium Salsa",
                 "Kettle Sweet Potato SeaSalt",
                 "Pringles SourCream  Onion",
                  "Doritos Corn Chips Original",
                 "Twisties Cheese Burger",
                 "Old El Paso Chunky Hot Tomato Salsa",
                 "Cobs Popd Sweet Chilli SourCream Chips",
                 "Woolworths Mild Salsa",
                 "Natural Chip Tomato Herbs Spices",
                 "Cobs Popd SeaSalt Chips",
                 "Smiths Crinkle Cheese Onion Chips",
                 "French Fries Potato Chips",
                 "Old el Paso Med Chunky Tomato Salsa",
                 "Doritos Corn Chips Cheese Supreme",
                 "Pringles Original Crisps",
                 "RRD Chilli Coconut",
                  "WW Original Corn Chips",
                  "Thins Hot Spicy Potato Chips",
                  "Cobs Popd Chips SourCREAM CHIVES",
                 "Smiths Crnkle Chip  Orgnl Big Bag",
                 "Dorito's Nacho Cheese Corn Chips",
                  "Kettle Sensations BBQ Maple",
                  "WW Deli Style Chips SeaSalt",
                 "Pringles Chicken Salt Crips",
                 "WW Original Stacked chips",
                  "Smiths Thinly Chips Salt Vinegar",
                  "Cheezels Cheese",
                  "Tostitos Lightly Salted",
                  "Thins Chips Salt Vinegar",
                 "Smiths Chinckle Chips Barbecue",
                 "Cheetos Puffs",
                  "RRD Sweet Chilli SourCream",
                 "WW Crincle Cut original",
                 "Tostitos Splash of Lime",
                 "Wooworths Medium Salsa",
                 "Kettle Tortilla Chips Beetroot Ricotta",
                 "CCs Tasty Cheese",
                 "Woolworths Cheese Rings",
                 "Tostitos Smoked Chipotle",
                 "Pringles Barbeque",
                 "WW Supreme Cheese Corn Chips",
                 "Pringles Mystery Flavour",
                 "Tyrells Crisps Cheddar Chives",
                 "Snbts Whlgrn Crisps CheddrMstrd",
                 "Cheetos Cheese Bacon Balls",
                 "Pringles Salt Vingar",
                 "Infuzions SourCream Herbs Veg",
                 "Kettle Tortilla ChpsFeta Garlic",
                 "Infuzions Papadams Mango Chutney",
                 "RRD Steak Chimichurri",
                 "RRD Honey Soy Chicken",
                 "Sunbites Whlegrn Crisps Frch Onin",
                 "RRD Salt Vinegar",
                 "Doritos Cheese Supreme",
                 "Smiths Crinkle Cut  SnagSauce ",
                 "WW Sour Cream OnionStacked Chips",
                 "RRD Lime Pepper",
                 "Natural Chip SeaSalt Vinegr",
                 "RRD Chikn Garlic Aioli",
                 "RRD SR Slow Rst Pork Belly",
                 "RRD Pc Sea Salt",
                 "Smith Crinkle Bolognese",
                 "Doritos Salsa Mild"))
```



```{r}
table(customers_transaction_dataset$PRODUCT_NAME)
```


```{r}
Numberextract <- function(string2){
  unlist(regmatches(string2,gregexpr("[[:digit:]]+\\.*[[:digit:]]*",string2)))
}
```


```{r}
customers_transaction_dataset["Package_Weight"] <- 
  Numberextract(customers_transaction_dataset$PROD_NAME)
```


```{r}
customers_transaction_dataset$Package_Weight <- 
  as.numeric(customers_transaction_dataset$Package_Weight)
```



Dropping Column
```{r}
customers_transaction_dataset02 <- customers_transaction_dataset %>% select(-c(PROD_NAME))
```


Finding duplicates rows 
```{r}
anyDuplicated(customers_transaction_dataset02)
```


```{r}
anyDuplicated(customers_transaction_dataset02$TXN_ID)
```


Selecting specific rows
```{r}
customers_transaction_dataset02[rownames(customers_transaction_dataset02) 
                               %in% c(40,41,42,43,124844,124845,124846),]
```


```{r}
filter(customers_transaction_dataset02,PROD_QTY >= 200) %>%
 select(STORE_NBR,LYLTY_CARD_NBR,TXN_ID,PROD_NBR,PROD_QTY) 
```


Deleting rows :
```{r}
customers_transaction_dataset02 <- customers_transaction_dataset02[
                              (customers_transaction_dataset02$PROD_QTY!=200),]
```


Removing duplicate rows in data frame:
```{r}
customers_transaction_dataset02 <- distinct(customers_transaction_dataset02)
```


```{r}
anyDuplicated(customers_purchase_dataset)
```


```{r}
which(customers_purchase_dataset$LYLTY_CARD_NBR==226000)
```

```{r}
customers_purchase_dataset[c(59693,59694,59695,59696,59697),]
```

Dropping row :
```{r}
customers_purchase_dataset <- customers_purchase_dataset %>%
                                        slice(-c(59695))
```



Merge keep all rows from both data frames:
 
```{r}
customers_dataset_merge <- merge(x=customers_transaction_dataset02,
                                y=customers_purchase_dataset,
                                by="LYLTY_CARD_NBR",
                                all=TRUE)
```


```{r}
sum(is.na(customers_dataset_merge))
```

```{r}
colnames(customers_dataset_merge)
```


```{r}
customers_dataset_merge <- customers_dataset_merge %>% 
                                relocate(
                                ï..DATE,PREMIUM_CUSTOMER,LIFESTAGE,LYLTY_CARD_NBR,
                                STORE_NBR,TXN_ID,PROD_NBR,PRODUCT_NAME,
                                Package_Weight,PROD_QTY,TOT_SALES)
```


```{r}
customers_dataset_merge$PREMIUM_CUSTOMER <- as.factor(customers_dataset_merge$PREMIUM_CUSTOMER)
```



```{r}
ggplot(data = customers_dataset_merge)+
  geom_bar(mapping = aes(x=LIFESTAGE, fill=PREMIUM_CUSTOMER)) +
  theme(legend.position = 'top')+
  coord_flip()
```



```{r}
sort(unique(customers_dataset_merge$TOT_SALES))
```



```{r}
ggplot(customers_dataset_merge, 
       aes(PROD_QTY, TOT_SALES)) + 
        geom_point()+
        geom_quantile() + 
        ggtitle(" Product Quantity vs Total Sales")+
        scale_x_discrete("PROD_QTY")+
        scale_y_continuous("TOT_SALES")
```


```{r}
ggplot(customers_dataset_merge, aes(PREMIUM_CUSTOMER,
                                    TOT_SALES)) + 
                                    geom_boxplot()
```


Selecting 500 samples and removing some columns :

```{r}
customers_dataset2_merge <- customers_dataset_merge[1:500,-c(1,2,3,8)]
```



#### Pre-processing for clustering

```{r}
 customers_dataset2_merge_scaled <- scale(customers_dataset2_merge)
```


```{r}
d <-  dist(x=customers_dataset2_merge_scaled, method ="euclidean")
```


```{r}
distance_matrix <- as.matrix(d)
```


```{r}
hc <-  fastcluster::hclust(d, method = "complete")
```



```{r}
plot(hc)
```


Cut into into 3 groups;
```{r}
sub_groups <- cutree(hc, k=3)
```


Number of members in  each cluster:
```{r}
table(sub_groups)
```


Drawing  the dendrogram with a border around the 3 clusters 

```{r}
plot(hc, cex = 0.9)
rect.hclust(hc, k=3, border = 2:4)
abline(h=3, col ="blue")
```


