---
title: "Experimentation and uplift testing"
author: "Jose Alberto"
date: "29/06/2021"
output: github_document
---
```{r}
rm(list=ls(all=TRUE))
```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load packages in r
```{r}
library(rmarkdown)
library(htmltools)
library(tidyverse)
library(dplyr)
library(plyr)
library(kableExtra)
library(stringr)
library(openxlsx)
library(lubridate)
library(ggplot2)
library(GGally)
library(fpp)
library(stats)
library(fastcluster)
```


```{r}
customers_chips_brand_dataset <- read.csv("datasets/QVI_data.csv")
```

The structure of the data
```{r}
str(customers_chips_brand_dataset)
```


```{r}
customers_chips_brand_dataset$DATE <- as.Date(customers_chips_brand_dataset$DATE)
```



```{r}
sum(is.na(customers_chips_brand_dataset))
```

```{r cars}
summary(customers_chips_brand_dataset)
```

```{r}
anyDuplicated(customers_chips_brand_dataset)
```

```{r}
customers_chips_brand_dataset <- customers_chips_brand_dataset %>%
  distinct()
```


```{r}
count_brand <- head(customers_chips_brand_dataset %>% 
  dplyr::count(BRAND),40)
```


```{r}
knitr::kable(count_brand)
```



Building a contigency table based on life stage and product name
```{r}
table(customers_chips_brand_dataset$LIFESTAGE, customers_chips_brand_dataset$PROD_NAME)
```

```{r}
chips_brand <- customers_chips_brand_dataset[
  customers_chips_brand_dataset$BRAND =="KETTLE"|
     customers_chips_brand_dataset$BRAND=="FRENCH"|
    customers_chips_brand_dataset$BRAND=="SMITHS"|
  customers_chips_brand_dataset$BRAND=="DORITOS"|
    customers_chips_brand_dataset$BRAND=="PRINGLES",]
```


```{r}
ggplot(data = chips_brand) + 
  geom_bar( mapping = aes(x= BRAND, fill=BRAND))
```


Density plot 
```{r}
ggplot(chips_brand, aes(x=TOT_SALES, fill=BRAND, colour=BRAND ))+
  geom_density(alpha = 0.5)
```

Density plot of the log10
```{r}
ggplot(chips_brand, aes(x=log10(TOT_SALES), fill=BRAND, colour=BRAND ))+
  geom_density(alpha = 0.5)
```



```{r}
customers_chips_brand_dataset2 <- customers_chips_brand_dataset %>% 
  select(PROD_QTY,TOT_SALES,BRAND)
```


Changing row names
```{r}
row.names(customers_chips_brand_dataset2) <- paste(customers_chips_brand_dataset2$BRAND, 1:nrow(customers_chips_brand_dataset2))
```


Removing column
```{r}
customers_chips_brand_dataset2 <- select(customers_chips_brand_dataset2, -BRAND)
```


Selecting 1000 samples

```{r}
customers_chips_brand_dataset2 <- 
  customers_chips_brand_dataset2 %>% slice(1:1000)
```


Standardization 
According to various studies , clustering with standardized attributes results into better cluster output compared to non standardized values
```{r}
customers_chips_brand_dataset2_scaled <- scale(customers_chips_brand_dataset2)
```



Implements hierarchical clustering with Euclidean distance
```{r}
hc.complete <- fastcluster::hclust(dist(customers_chips_brand_dataset2_scaled),
                                   method="complete")
```



```{r}

plot(hc.complete, main = "Complete Linkage", xlab = " Chips Brand", cex = 1)
```

```{r}
sub_groups <- cutree(hc.complete, 3)
```


```{r}
table(sub_groups)
```
